{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion}

### 2017-01-04 proposal/quicksand.tex
quicksand eval

% TODO: is related work needed?
% i can easily talk about it here but i dont' want to set a weird precedent for super weak related work in my other sections.

>>>>

I evaluated Quicksand on the product of the following two questions: For an arbitrary fixed CPU budget...

\begin{itemize}
	\item ...does Quicksand find more bugs than a single-state-space (SSS-MC) approach...
	\item ...does Quicksand provide more total verifications than SSS-MC...
\end{itemize}
...where the SSS-MC control is configured to preempt...
\begin{itemize}
	\item ...on synchronization API boundaries only?
	\item ...on every single shared memory access?
\end{itemize}

I also conducted a separate experiment focused on the issue of {\em nondeterministic} data-races:
candidates that require model checking to expose in the first place, which would be
false-negatives during a single-execution analysis.

I also also ran several different configurations of Quicksand in each of these experiments,
alternately using Limited Happens-Before or Pure Happens-Before for its data-race analysis.
Both of these cases outperformed all the other cases of SSS-MC, so this was mostly useful for comparing the two techniques against each other.

\subsection{Setup and Test Suite}

The test suite for all these experiments was 157 student projects: 79 P2 thread libraries from CMU, and 78 Pintos kernels from Berkeley and the University of Chicago (Section~\ref{sec:overview-edu}).
I paired each submission with several test cases, 6 for each P2 and 3 for each Pintos.
In total there were 629 unique pairs of a test case and a student project, henceforth called simply ``tests''.
(Savvy mathematicians will note that doesn't quite add up -- some of the Pintoses were incomplete, and could only run 1 or 2 tests.)

For each of the 629 tests, I ran Quicksand and SSS-MC for 10 CPU-hours each in each of the configurations named above.
Quicksand, being inherently parallel, was given 10 CPUs for 1 wall-clock hour,
while SSS-MC control experiments were given 10 wall-clock hours on 1 CPU
(which is being charitable: comparing CPU-time gives the controls perfectly efficient hypothetical scaling from parallelization).

\subsection{Results}

The results are shown in Figure~\ref{fig:graphs}.
The first 3 graphs all plot a cumulative distribution of their respective achievements,
with elapsed time on the X-axis (as a log scale, to emphasize detail in the ``easy'' tests).
The captions largely explain each result.
To summarize, Quicksand outperforms the state-of-the-art on all metrics, excepting the impact of parallelization start-up overhead for very small CPU-time limits.
The Limited HB data-race strategy generally finds bugs faster when they exist,
but suffers on total verifications compared to Pure HB,
because of its false positives which create excess redundant state spaces for Landslide to check.

\begin{figure}[h]
	\begin{tabular}{cc}
		\begin{tabular}{p{0.48\textwidth}p{0.48\textwidth}}
			\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-v2.pdf}
		&
			\includegraphics[width=0.48\textwidth]{dowefindbugsfaster-wallclock-v2.pdf}
		\\
			(a)
			Bugs found as a function of elapsed CPU time. After the break-even point at
			\textasciitilde{}200 seconds, Quicksand outperforms all control experiments.
		&
			(b)
			Bugs found by elapsed wall-clock time.
			This presentation emphasizes that the ``break-even'' point in (a)
			is just an artifact of parallelization start-up overhead.
			Quicksand is parallelized tenfold; the vertical line indicates its 1 hour limit.
		\\
			\includegraphics[width=0.48\textwidth]{totalverifs-v2.pdf}
		&
			\includegraphics[width=0.48\textwidth]{nondets.pdf}
		\\
			(c)
			Total verifications provided as a function of CPU time.
			The lines for SSS-MC-ICB and SSS-MC-DPOR are artificially penalized to exclude tests with data races,
			as preempting on sync APIs alone is not sufficient to verify those.
			SSS-MC-Shared-Mem, however, is not penalized thus;
			its failure here is due to the massive computational overhead from so many preemption points.
		&
			(d)
			Plot of how many data-race bugs were found using ``single-pass'' data-race candidates versus ``nondeterministic'' ones.
			With both Limited and Pure HB approaches, the MC's ability to find new data-race candidates in obscure thread interleavings led to more bugs found in all test cases.
			Between Limited and Pure HB, Pure HB is more reliant on this phenomenon, as Limited HB can often find more potential races on the first pass.
		\end{tabular}
	\end{tabular}
	\caption{Results.}
	\label{fig:graphs}
\end{figure}
